---
title: "Playing with ggplot2"
author: "Raju Rimal"
date: "16 Nov 2014"
output:
  html_document:
    keep_md: yes
---


Diagnostic Plots
----------

I have tried to use fortify function in ```ggplot2``` which can access different statistics related to linear model. The basic diagnostic plot which we often get using plot function in the fitted model using ```lm``` command.

The function ```diagPlots``` gives an list of six different plots which can be arranged in a grid using ```grid``` and ```gridExtra``` packages.
```{r results='hide', message=FALSE, warning=FALSE}
require(ggplot2)
diagPlot<-function(model){
    p1<-ggplot(model, aes(.fitted, .resid))+geom_point()
    p1<-p1+stat_smooth(method="loess")+geom_hline(yintercept=0, col="red", linetype="dashed")
    p1<-p1+xlab("Fitted values")+ylab("Residuals")
    p1<-p1+ggtitle("Residual vs Fitted Plot")+theme_bw()
    
    p2<-ggplot(model, aes(qqnorm(.stdresid)[[1]], .stdresid))+geom_point(na.rm = TRUE)
    p2<-p2+geom_abline(aes(qqline(.stdresid)))+xlab("Theoretical Quantiles")+ylab("Standardized Residuals")
    p2<-p2+ggtitle("Normal Q-Q")+theme_bw()
    
    p3<-ggplot(model, aes(.fitted, sqrt(abs(.stdresid))))+geom_point(na.rm=TRUE)
    p3<-p3+stat_smooth(method="loess", na.rm = TRUE)+xlab("Fitted Value")
    p3<-p3+ylab(expression(sqrt("|Standardized residuals|")))
    p3<-p3+ggtitle("Scale-Location")+theme_bw()
    
    p4<-ggplot(model, aes(seq_along(.cooksd), .cooksd))+geom_bar(stat="identity", position="identity")
    p4<-p4+xlab("Obs. Number")+ylab("Cook's distance")
    p4<-p4+ggtitle("Cook's distance")+theme_bw()
    
    p5<-ggplot(model, aes(.hat, .stdresid))+geom_point(aes(size=.cooksd), na.rm=TRUE)
    p5<-p5+stat_smooth(method="loess", na.rm=TRUE)
    p5<-p5+xlab("Leverage")+ylab("Standardized Residuals")
    p5<-p5+ggtitle("Residual vs Leverage Plot")
    p5<-p5+scale_size_continuous("Cook's Distance", range=c(1,5))
    p5<-p5+theme_bw()+theme(legend.position="bottom")
    
    p6<-ggplot(model, aes(.hat, .cooksd))+geom_point(na.rm=TRUE)+stat_smooth(method="loess", na.rm=TRUE)
    p6<-p6+xlab("Leverage hii")+ylab("Cook's Distance")
    p6<-p6+ggtitle("Cook's dist vs Leverage hii/(1-hii)")
    p6<-p6+geom_abline(slope=seq(0,3,0.5), color="gray", linetype="dashed")
    p6<-p6+theme_bw()
    
    return(list(rvfPlot=p1, qqPlot=p2, sclLocPlot=p3, cdPlot=p4, rvlevPlot=p5, cvlPlot=p6))
}
```

Using the ```mtcars``` datasets, a linear model is fitted with mpg as response and cyl, disp, hp, drat and wt has predictor variable

```{r}
lm.model<-lm(mpg~cyl+disp+hp+drat+wt, data=mtcars)
diagPlts<-diagPlot(lm.model)
```

To display the plots in a grid, some packages mentioned above should be installed.

```{r, results='hide', warning=FALSE, message=FALSE}
lbry<-c("grid", "gridExtra")
lapply(lbry, require, character.only=TRUE, warn.conflicts = FALSE, quietly = TRUE)
```
Thus the plot obtained is,

```{r}
do.call(grid.arrange, c(diagPlts, main="Diagnostic Plots", ncol=3))
```